From DSL to HPC Component-Based Runtime:\\ A Multi-Stencil DSL Case Study

### 2015-08-21 WOLFHPC15/article/ms_msc.tex
Rel S4

\subsubsection*{Computation Domains and Their Dependencies}
A computation domain is a subset of a given kind of mesh entities, used for at least one computation.
For each computation domain, the kind of mesh entities on which it is mapped is specified (lines 3-4 and 12-14 on Figure~\ref{fig:grammar}).
By default, two computation domains of the same entities are handled as intersecting except when an explicit independence relation is mentioned in the \emph{independent} section of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
On lines 4 to 6 of Figure~\ref{fig:mslex}, computation domains are defined for each kind of mesh entities: \texttt{allcell}, \texttt{alledgex} and \texttt{alledgey}; they are used to represent the whole domain.
On lines 7 and 8 two other domains of \texttt{edgex} are defined: \texttt{part1edgex} and \texttt{part2edgex}; they are used to represent the boundaries of the domain.
On line 10 the independence relation between \texttt{part1edgex} and \texttt{part2edgex} is specified.

@@

\subsubsection*{Computation Description}
The last part of the language contains the specifications of stencil and local kernels as defined in Section~\ref{sect:multistencil}.
This description does not contain a direct expression of the numerical expressions $exp$.
Instead, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} simply corresponds to an identifier that references an implementation in an external language  (\cf Section~\ref{sect:component}).
The description starts by the identifier of the computed data element ($w$ in the formal description) with the computation domain between brackets ($d$).
Then, after the equal sign, the kernel identifier ($exp$) is specified.
Finally, between the parenthesis is the list of identifiers of the data elements read by the computation with their stencil shape between brackets ($R$).
If the computation is local, no brackets appear ($R_l$). In the example of Figure~\ref{fig:mslex}, nine computations are described. On line 24, the data element \texttt{b} is computed on the domain \texttt{allcell} by the kernel expression \texttt{c\_0} which read data \texttt{a} without neighborhood shape.

%----------------------------------------
\subsection{MSC Parallelization}
\label{sect:msc}
%----------------------------------------
The MSC parallelization is a subpart of the overall compiler MSCAC.
It makes use of the ordered list of computations $\Gamma$, which can directly be extracted from the parser, to build a parallel representation of the computations of the overall multi-stencil program.
This parallelization phase is divided in five different steps to transform $\Gamma$ to a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393}. These steps are described in this section.
%This section gives an overview of these five steps , but they are more formally defined with their associated algorithms in a research report~\cite{????}.

@@

\subsubsection*{The Synchronized Ordered List of Computations: $\Gamma_{sync}$}
In a data parallelization, data is split among processors and the same program is applied on each sub-part of the data.
However, this parallelization technique requires additional synchronizations between processors.
%For example, as a stencil kernel accesses neighbor values, values computed by another processor are needed and must be synchronized. 
The required synchronizations can be automatically computed from the ordered list of computations $\Gamma$.
A synchronization is needed each time a data read by a stencil computation has been written by a previous computation. In such a case, a computation is added before the stencil computation. This \emph{synchronization computation} reads the data to synchronize, and write the same data. The computation domain of such a synchronization computation encompasses the whole mesh entities on which the data is declared. As a result, $\Gamma$ is transformed to a synchronized ordered list of computations $\Gamma_{sync}$.
For example, in Figure~\ref{fig:mslex}, the stencil computation $c_1$ reads the data element $b$ which has been written by $c_0$. For this reason the sublist $[c_0,c_1]$ of $\Gamma$ is transformed into $[c_0,sync_1,c_1]$ in $\Gamma_{sync}$. The new computation $sync_1$ reads and writes $b$ and is applied on \texttt{edgex}. As a result, a dependency is kept between $c_0$ and $sync_1$ and between $sync_1$, and $c_1$. The same transformation is performed for the two other stencils: $c_4$ and $c_8$. Thus, $\Gamma_{sync} = [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$.

@@

  \path[->]
    (c0) edge node {} (sy1)
    (sy1) edge node {} (c1)
    (c1)  edge node {} (c2)
          edge node {} (c3)
    (c2) edge node {} (sy4)
    (sy4) edge node {} (c4)
    (c4) edge node {} (c6)
    (c3) edge node {} (c5)
    (c5) edge node {} (c7)
    (c6) edge node {} (c7)
    (c7) edge node {} (sy8)
    (sy8) edge node {} (c8);
\end{tikzpicture}
\caption{$\Gamma_{hybrid}$ of the example of Figure~\ref{fig:mslex}}
\label{fig:hyb}
\end{center}
\end{figure}

@@

\begin{figure}[h!]
\begin{center}
\subfloat[][The forbidden $N$ shape.\label{fig:n}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (a) at (0,0) {$c_0$};
   \node[] (b) at (2,0) {$c_1$};
   \node[] (c) at (0,-1) {$c_2$};
   \node[] (d) at (2,-1) {$c_3$};

>>>>

\subsubsection*{Computation Domains and Their Dependencies}
A computation domain is a subset of a given kind of mesh entities, used for at least one computation.
For each computation domain, the kind of mesh entities on which it is mapped is specified (lines 3-4 and 12-14 on Figure~\ref{fig:grammar}).
By default, two computation domains of the same entities are handled as intersecting except when an explicit independence relation is mentioned in the \emph{independent} section of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
On lines 4 to 6 of Figure~\ref{fig:mslex}, computation domains are defined for each kind of mesh entities: \texttt{allcell}, \texttt{alledgex} and \texttt{alledgey}; they are used to represent the whole domain.
On lines 7 and 8, two other domains of \texttt{edgex} are defined: \texttt{part1edgex} and \texttt{part2edgex}; they are used to represent the boundaries of the domain.
On line 10 the independence relation between \texttt{part1edgex} and \texttt{part2edgex} is specified.

@@

\subsubsection*{Computation Description}
The last part of the language contains the specifications of stencil and local kernels as defined in Section~\ref{sect:multistencil}.
This description does not contain a direct expression of the numerical expressions $exp$.
Instead, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} corresponds to an identifier that references an implementation in an external language  (\cf Section~\ref{sect:component}).
The description starts by the identifier of the computed data element ($w$ in the formal description) with the computation domain between brackets ($d$).
Then, after the equal sign, the kernel identifier ($exp$) is specified.
Finally, between the parenthesis is the list of identifiers of the data elements read by the computation with their stencil shape between brackets ($R$).
If the computation is local, no brackets appear ($R_l$). In the example of Figure~\ref{fig:mslex}, nine computations are described. On line 24, the data element \texttt{b} is computed on the domain \texttt{allcell} by the kernel expression \texttt{c\_0} which reads data \texttt{a} without any neighborhood shape.

%----------------------------------------
\subsection{MSC Parallelization}
\label{sect:msc}
%----------------------------------------
The MSC parallelization is a subpart of the overall compiler MSCAC.
It makes use of the ordered list of computations $\Gamma$, which can directly be extracted from the parser, to build a parallel representation of the computations of the overall multi-stencil program.
This parallelization phase is divided in five different steps to transform $\Gamma$ to a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393}. This sections describes these steps.
%This section gives an overview of these five steps , but they are more formally defined with their associated algorithms in a research report~\cite{????}.

@@

\subsubsection*{The Synchronized Ordered List of Computations: $\Gamma_{sync}$}
Data parallelization splits data among processors and the same program is applied on each sub-part of the data.
However, this parallelization technique requires additional synchronizations between processors.
%For example, as a stencil kernel accesses neighbor values, values computed by another processor are needed and must be synchronized. 
The required synchronizations can be automatically computed from the ordered list of computations $\Gamma$.
A synchronization is needed each time a data read by a stencil computation has been written by a previous computation. In such a case, a computation is added before the stencil computation. This \emph{synchronization computation} reads the data to synchronize, and write the same data. The computation domain of such a synchronization computation encompasses the whole mesh entities on which the data is declared. As a result, $\Gamma$ is transformed to a synchronized ordered list of computations $\Gamma_{sync}$.
For example, in Figure~\ref{fig:mslex}, the stencil computation $c_1$ reads the data element $b$ which has been written by $c_0$. For this reason the sublist $[c_0,c_1]$ of $\Gamma$ is transformed into $[c_0,sync_1,c_1]$ in $\Gamma_{sync}$. The new computation $sync_1$ reads and writes $b$ and is applied on \texttt{edgex}. As a result, a dependency is kept between $c_0$ and $sync_1$ and between $sync_1$, and $c_1$. The same transformation is performed for the two other stencils: $c_4$ and $c_8$. Thus, for the example of Figure~\ref{fig:mslex}, $\Gamma_{sync} = [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$.

@@

  \path[->]
    (c0) edge node {} (sy1)
    (sy1) edge node {} (c1)
    (c1)  edge node {} (c2)
          edge node {} (c3)
    (c2) edge node {} (sy4)
    (sy4) edge node {} (c4)
    (c4) edge node {} (c6)
    (c3) edge node {} (c5)
    (c5) edge node {} (c7)
    (c6) edge node {} (c7)
    (c7) edge node {} (sy8)
    (sy8) edge node {} (c8);
\end{tikzpicture}
\caption{$\Gamma_{hybrid}$ of the example of Figure~\ref{fig:mslex}.}
\label{fig:hyb}
\end{center}
\end{figure}

@@

\begin{figure}[h!]
\begin{center}
\subfloat[][The forbidden $N$ shape.\\\label{fig:n}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (a) at (0,0) {$c_0$};
   \node[] (b) at (2,0) {$c_1$};
   \node[] (c) at (0,-1) {$c_2$};
   \node[] (d) at (2,-1) {$c_3$};
